ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-09 rebuttal.tex

\begin{reviewer}
    %
    2.~Am I correct to state that ChemKED is a YAML file with a predefined structure?
    %
\end{reviewer}

\begin{response}
    %
    Yes, ChemKED files are files in YAML format, with a prescribed structure and set of contents.
    The required structure and contents are defined by our schema, which is documented in this
    manuscript and in the online documentation.
    %
\end{response}

\begin{reviewer}
    %
    3.~I think the article would be more appealing if it would have at least one figure as an
    example.
    %
\end{reviewer}

\begin{response}
    In response to one of your later comments, we have extended our examples to make them
    more usable as-is, and also added figures to show results from these examples on
    \textbf{pages 16 and 19}.
\end{response}

@@

    Moreover, all existing databases have a similar requirement (whether explicitly stated or
    not), where users submit data files for inclusion that are approved by the database owners after
    some review. In this sense, we actually see the Pull Request mechanism of GitHub as an
    advantage, since all reviews can be done publicly and some of the review procedures (style
    checks, reference validity, etc.) may be done automatically by a script associated with the
    GitHub repository. In this way, the entire community can contribute to improving the quality of
    the database.

    Lastly, while we hope other researchers will contribute to this database (such that it becomes a
    true community effort), there is no requirement for others to use our database to use ChemKED
    files or PyKED. Others are free to use the tools internally, or download\slash fork our database
    and build their own. Moreover, as mentioned in the manuscript, we have included the ability to
    generate ReSpecTh-compliant files from ChemKED files, so that the community can move data to
    alternative formats easily and so there is a reduced chance of a negative ``lock-in'' effect.
    %
\end{response}

\begin{reviewer}
    %
    14.~Page 6, line 18 states that a converter exists from ReSpecTh to ChemKED. But isn't this
    feature added when going from v0.1.6 to v0.2.0, which makes page 3 line 21: v0.1.6 not correct.
    %
\end{reviewer}

\begin{response}
    %
    In principle, the version of the ChemKED database format and the version of the PyKED Python
    package are entirely independent. Thus, since adding the converters incurred no changes to the
    database format, the description of the format in the manuscript is a description of v0.1.6,
    which happens to be identical to v0.2.0. However, in practice, the versions of PyKED and the
    ChemKED database have been coupled up to now, and we have updated the version mentioned in the
    text on \textbf{page 3} to v0.3.0, which we released in conjunction with submitting
    this response after the helpful comments of the reviewers.
    %
\end{response}

### 2017-10-09 rebuttal.tex

% bump up section font size
\usepackage{sectsty}
\sectionfont{\fontsize{12}{15}\selectfont}

\usepackage[colorlinks=true]{hyperref}
\usepackage{todonotes}

\setlength{\parindent}{0in}
\setlength{\parskip}{0.25cm}

\newenvironment{reviewer}{\vspace{0.5\baselineskip}\begingroup\itshape\textbf{Reviewer:}}{\endgroup}
\newenvironment{response}{\textbf{Our response:}}{\vspace{0.5\baselineskip}}
\newcommand{\review}[1]{{\itshape#1}}
\newcommand{\respond}[1]{\textbf{Our response:} #1}

@@

\section*{Reviewer \#1:}

\begin{reviewer}
    %
    Currently four database formats are offered for storing combustion experimental data. These are
    CloudFlame (Excel/CSV format), PrIMe (XML-based format), ReSpecTh (also XML-based format), and
    ChemKED (YAML-based format). Elaboration of the last format is reported in the submitted
    manuscript. The motivation for the elaboration of this new format was based on the statement
    that YAML is more human readable than XML. I think that this is simply not true. Even a
    first-time user can read the raw datafiles of both formats with equal efficiency. Also,
    currently the suggested new format is much more limited (e.g. it is able to encode ignition
    delay time experiments only). The presently planned features of ChemKED are already available in
    the other three formats.

    However, I suggest the publication of an improved and shortened version of the manuscript in
    IJCK. The reason is that a permanent and laboratory-independent storage of experimental data is
    an important issue and I agree that a public discussion promotes the acceptance of a widely used
    common format.
    %
\end{reviewer}

\begin{response}
    %
    We thank the reviewer for their comments. Detailed responses to each comment are
    below.

    Regarding the manuscript, please see our response to your next point; in addition, we added more
    detail to our examples in response to Reviewer \#2's comments, which increased the length of
    the manuscript slightly.
    %
\end{response}

@@

    On the second point, we agree with the reviewer that CSV or Excel files may give an adequate
    summary or overview of the data, but only if the files are opened with some appropriate
    software, e.g., Microsoft Excel (particularly for Excel-formatted files). Thus, in this sense,
    CSV or Excel files do not offer any advantage over ChemKED files (or any other database files)
    that must also be processed with an appropriate software package (PyKED, in the case of ChemKED
    files) so that a summary can be produced. Furthermore, in the specific case of PyKED/ChemKED, it
    is already possible to produce an Excel or CSV file (using the Pandas DataFrame integration) so
    that one may view such a summary if desired. See also the response to a similar comment below.
    %
\end{response}

\begin{reviewer}
    %
    1.~How are the authors planning on `forcing' the experimentalists to publish their data in this
    new ChemKED file format?
    %
\end{reviewer}

\begin{response}
    %
    While we cannot force anyone to use a particular format, we hope that by defining a format and
    ecosystem with desirable features---including inherent openness and usability---experimentalists
    will begin adopting a community-standard format. In fact, we have already received expressions
    of interest from various experimental groups, and hope this will grow as we continue to improve
    and add features. We also hope that at a certain point, with wide adoption, sharing experimental
    data in a standard format will become the expected norm (e.g., when submitting an article for
    publication) like sharing a Chemkin-format chemical kinetic model.

    Furthermore, we are willing to work (and are currently working) directly with experimentalists
    to provide them easy-to-use conversion scripts from internal data formats. In return, they are
    helping us define what data and information needs to go in the specification for their
    experiments.
    %
\end{response}

@@

\begin{response}
    %
    The reviewer is correct, and this was a typo in the abstract. We have clarified that PyKED is
    intended as software to validate and work with ChemKED files.

    That said, we do have ongoing work to create a graphical user interface to create ChemKED files,
    as this was requested by several of the experimental teams that have expressed interest in the
    ChemKED format. The work is available publicly at
    \url{https://github.com/pr-omethe-us/ChemKED-gui}. In addition, we have a built-in converter
    for generating ChemKED files from existing ReSpecTh files.
    %
\end{response}

\begin{reviewer}
    %
    5.~Page 1, line 31: The future development plans include support for jet stirred reactors,
    however, this is not mentioned in section 5, Conclusions and future work, nor is there any
    reference to it on the GitHub wiki roadmap of PyKED.
    %
\end{reviewer}

\begin{response}
    %
    Thank you for the comment, we have corrected this oversight in the roadmap and the
    conclusions (\textbf{page 20}).
    %
\end{response}

@@

\begin{reviewer}
    %
    23.~Page 10, line 29--30: ``A user might want to load the information'' Avoid the term `might'.
    It's one of the main goals of using ChemKED and its associated Python-package PyKED to load the
    information necessary for simulating the experiment.
    %
\end{reviewer}

\begin{response}
    %
    We have modified the text on \textbf{page 12} to describe this example as an application.
    %
\end{response}

\begin{reviewer}
    %
    24.~The Python code snippets are not always initiated in the text.
    %
\end{reviewer}

\begin{response}
    %
    We have corrected this oversight, with more explanation and introduction of the Python
    examples on \textbf{pages 12--19}.
    %
\end{response}

@@

\begin{response}
    %
    \verb|air.xml| is an input file included with Cantera, representing an average composition of
    air. We have modified the text on \textbf{page 13} to include this information.
    %
\end{response}

\subsubsection*{Shock tube modelling with constant volume}

\begin{reviewer}
    %
    27.~The table overlapping page 11 and 12 can be reduced by truncating the \texttt{DataPoints},
    similarly as is done with the table starting at page 9, line 31 and ending at page 10, line 27.
    %
\end{reviewer}

\begin{response}
    %
    We have truncated the datapoints list on \textbf{pages 12 and 17} as the reviewer suggests.
    %
\end{response}

@@

\begin{reviewer}
    %
    28. Python code snippet on page 13: (a.) Add more commentary text. (b.) Changing the example to
    enable multithreading might make it more appealing to start using ChemKED as a data format in
    addition with Cantera as a simulation tool.
    %
\end{reviewer}

\begin{response}
    %
    We have modified the example on \textbf{pages 17--19} to include more
    explanatory text and to use Python's built-in
    multiprocessing library to run the simulations on many processors. We note that if Cantera has
    been built with multi-threaded BLAS and LAPACK libraries (such as the Intel MKL), then the 0-D
    solver will automatically utilize the multithreading capabilities of those libraries. Therefore,
    using the multiprocessing library may not give much benefit in terms of overall speedup,
    depending on the complexity of the chemical kinetic model (i.e., how long the simulation
    takes to run), how many processors are available, and the availability of the multi-threaded
    BLAS and LAPACK libraries. However, we have not explained this detail of Cantera in the text to
    avoid confusing the reader.
    %
\end{response}

@@

\subsection*{Conclusions and future work}

\begin{reviewer}
    %
    29.~No conclusion is mentioned for the usage examples.
    %
\end{reviewer}

\begin{response}
    %
    We have added a paragraph to the conclusion on \textbf{page 20}.
    %
\end{response}

### 2017-10-10 pyked.tex
updated PyKED version in paper

This section provides an introduction to the current version of the \ck{} database format, v0.1.6,
which represents a functional standard for documenting autoignition measurements in shock tubes and
rapid compression machines. Here we describe the fields that a user creating a \ck{} file is likely
to use. There are several fields that are used internally in the schema that are not described here,
but can be found in the online documentation for the
schema.\footnote{\url{https://pr-omethe-us.github.io/PyKED/schema-docs.html}} In
Section~\ref{sec:usage-example}, we provide examples of \ck{} files and representative use cases. We
indicate \yabox|yaml| keywords or values using text with gray background, and \pybox|Python| code
using text with a yellow background.

>>>>

This section provides an introduction to the current version of the \ck{} database format, v0.3.0,
which represents a functional standard for documenting autoignition measurements in shock tubes and
rapid compression machines. Here we describe the fields that a user creating a \ck{} file is likely
to use. There are several fields that are used internally in the schema that are not described here,
but can be found in the online documentation for the
schema.\footnote{\url{https://pr-omethe-us.github.io/PyKED/schema-docs.html}} In
Section~\ref{sec:usage-example}, we provide examples of \ck{} files and representative use cases. We
indicate \yabox|yaml| keywords or values using text with gray background, and \pybox|Python| code
using text with a yellow background.
